---
title: "WGCNA_Cvriginica_Feeding_VennModules"
author: "Samuel Gurr"
date: "2023-08-31"
output: html_document
---

# SETUP
```{r setup, include= FALSE}
# LOAD PACKAGES
library(dplyr)
library(ggplot2)
library(dplyr)
library(VennDiagram)
library(ggVennDiagram)
library(ggvenn)
library(gridExtra)
# SET WORKING DIRECTORY

knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/Cvirginica_KM_paper/RAnalysis") # sets the working 

```


# LOAD THE MODULE-MEMBERSHIP MASTER TABLES 
```{r} 

Day0_ModuleMembership  <- read.csv(file="Output/Day0/Day0.WGCNA_ModulMembership.csv", sep=',', header=TRUE) 

Day1_ModuleMembership  <- read.csv(file="Output/Day1/Day1.WGCNA_ModulMembership.csv", sep=',', header=TRUE) 
```

## About:

Modules of interest 

Day 0 (at the end of the starvation period)

High expression Fed = turquoise

High expression Starved = blue 

High expression Wild = yellow 

High expression Lola = black 


Day 1 (after several days of recovery)

High expression Fed = greenyellow  **compare to Turquoise day 0**

High expression Starved = green  **compare to Blue day 0**

High expression Wild =  brown **comapre to Yellow day 0**

High expression Lola = blue - **compare to Black day 0**

# VENN DIAGRAMS - DATA PREP
```{r} 
Day0_Starved <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'turquoise') %>% dplyr::select('geneSymbol') %>% na.omit()
Day0_Fed     <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'blue') %>% dplyr::select('geneSymbol') %>% na.omit()
Day0_Wild    <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'yellow') %>% dplyr::select('geneSymbol') %>% na.omit()
Day0_Lola    <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'black') %>% dplyr::select('geneSymbol') %>% na.omit()


Day1_Starved <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'green') %>% dplyr::select('geneSymbol') %>% na.omit()
Day1_Fed     <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'greenyellow') %>% dplyr::select('geneSymbol') %>% na.omit()
Day1_Wild    <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'brown') %>% dplyr::select('geneSymbol') %>% na.omit()
Day1_Lola    <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'blue') %>% dplyr::select('geneSymbol') %>% na.omit()

```


# VENN DIAGRAMS - FEEDING TREATMENT
```{r}
Feeding_treatment_all <- list(
  Fed_d0      = Day0_Fed$geneSymbol, 
  Fed_d1      = Day1_Fed$geneSymbol, 
  Starved_d0  = Day0_Starved$geneSymbol,
  Starved_d1  = Day1_Starved$geneSymbol
)
FeedTreat_Venn_All <- ggvenn(Feeding_treatment_all, 
                   fill_color = c("#999999", "#E69F00", "#56B4E9", "#009E73"),
                   stroke_size = 0.5, set_name_size = 4)
# starved
Starved <- list(
  Starved_d0  = Day0_Starved$geneSymbol,
  Starved_d1  = Day1_Starved$geneSymbol
)
FeedTreat_Venn_Starved <- ggvenn(Starved, 
                   fill_color = c("#999999", "#009E73"),
                   stroke_size = 0.5, set_name_size = 4)
# fed
Fed <- list(
  Fed_d0      = Day0_Fed$geneSymbol, 
  Fed_d1      = Day1_Fed$geneSymbol
)
FeedTreat_Venn_Fed <- ggvenn(Fed, 
                   fill_color = c("#E69F00", "#56B4E9"),
                   stroke_size = 0.5, set_name_size = 4)

# merge fed and starved
FedStarved_Venns <- ggarrange(FeedTreat_Venn_Fed,
                              FeedTreat_Venn_Starved, nrow = 1)
# full figure 
Venn_all <- grid.arrange(FeedTreat_Venn_All, 
                         FedStarved_Venns,  nrow=2, clip="off")

#print
pdf(paste("Output/WGCNAMods_VennFeeding.pdf", sep =''), 
    width=7, 
    height=7)
print(FeedTreat_Venn_All)
dev.off()
          

```



# VENN DIAGRAMS - Line
```{r}
Line_all <- list(
  Wild_d0      = Day0_Wild$geneSymbol, 
  Wild_d1      = Day1_Wild$geneSymbol, 
  Lola_d0      = Day0_Lola$geneSymbol,
  Lola_d1      = Day1_Lola$geneSymbol
)
Line_Venn_All <- ggvenn(Line_all, 
                   fill_color = c("#999999", "#E69F00", "#56B4E9", "#009E73"),
                   stroke_size = 0.5, set_name_size = 4)
# Wild
Wild <- list(
  Wild_d0  = Day0_Wild$geneSymbol,
  Wild_d1  = Day1_Wild$geneSymbol
)
Line_Venn_Wild <- ggvenn(Wild, 
                   fill_color = c("#999999", "#009E73"),
                   stroke_size = 0.5, set_name_size = 4)
# Lola
Lola <- list(
  Lola_d0      = Day0_Lola$geneSymbol, 
  Lola_d1      = Day1_Lola$geneSymbol
)
Line_Venn_Lola <- ggvenn(Lola, 
                   fill_color = c("#E69F00", "#56B4E9"),
                   stroke_size = 0.5, set_name_size = 4)

# merge 
Line_Venns <- ggarrange(Line_Venn_Wild, 
                        Line_Venn_Lola, nrow = 1)
# full figure 
Line_Venn_all <- grid.arrange(Line_Venn_All, 
                              Line_Venns,  nrow=2, clip="off")

# print
pdf(paste("Output/WGCNAMods_VennLine.pdf", sep =''), 
    width=7, 
    height=7)
print(Line_Venn_All)
dev.off()
```


