---
title: "WGCNA_Cvirginica_KEGG"
author: "Samuel Gurr"
date: "2023-08-31"
output: html_document
---

```{r setup, include=FALSE}
# library(reactome.db)
library(clusterProfiler)
library(KEGGREST)
library(tidyr)
library(stringr)
library(forcats)
library(ggplot2)
library(scales)
library(ape)
library(data.table)
library(tidyverse)
library(fBasics)
library(dplyr)
library(ggplot2)
# SET WORKING DIRECTORY
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/Cvirginica_KM_paper/RAnalysis") # sets the working 
```


# LOAD THE MODULE-MEMBERSHIP MASTER TABLES 
```{r} 

Day0_ModuleMembership  <- read.csv(file="Output/Day0/Day0.WGCNA_ModulMembership.csv", sep=',', header=TRUE) 

Day1_ModuleMembership  <- read.csv(file="Output/Day1/Day1.WGCNA_ModulMembership.csv", sep=',', header=TRUE) 
```

# Prepare Cgigas KEGG IDs
```{r}

# Using KEGGREST..  KEGGREST prep 
pathways.list <- keggList("pathway", "crg")
head(pathways.list)
# Pull all genes for each pathway
pathway.codes <- sub("path:", "", names(pathways.list)) 
genes.by.pathway <- sapply(pathway.codes,
                           function(pwid){
                             pw <- keggGet(pwid)
                             if (is.null(pw[[1]]$GENE)) return(NA)
                             pw2 <- pw[[1]]$GENE[c(TRUE,FALSE)] # may need to modify this to c(FALSE, TRUE) for other organisms
                             pw2 <- unlist(lapply(strsplit(pw2, split = ";", fixed = T), function(x)x[1]))
                             return(pw2)
                           }
)

# prep for rich factor - we need the number of genes POSSIBLE per pathway
outdat   <- data.frame() # start dataframe 
outdat   <- data.frame(matrix(nrow = 1, ncol = 2)) # create dataframe to save cumunalitively during for loop
colnames(outdat) <- c('KEGGID_pathway', 'Num.genes.per.pathway')
Genes_per_pathway <- data.frame()
 for (i in 1:nrow(pVals.by.pathway)) {
  outdat$KEGGID_pathway        <- rownames(as.data.frame(pVals.by.pathway)[i,])
  outdat$Num.genes.per.pathway <- nrow(as.data.frame(genes.by.pathway[outdat$KEGGID_pathway[1]]))
  
  Genes_per_pathway <- rbind(Genes_per_pathway,outdat) #bind to a cumulative list dataframe
  #print(Genes_per_pathway) # print to monitor progress 
 }
Genes_per_pathway <- Genes_per_pathway %>% # ommit NAs
  dplyr::mutate(Num.genes.per.pathway = 
                  case_when(Num.genes.per.pathway == 1 ~ 0, # NAs are registered as N=1, make these 0
                            TRUE ~ Num.genes.per.pathway)) # all other cases simply print 

nrow(Genes_per_pathway) == 135 # must be TRUE

```

# DAY 0 DATA - FOR LOOP 

## About:

Modules of interest 

Day 0 (at the end of the starvation period)

High expression Fed = turquoise

High expression Starved = blue 

High expression Wild = yellow 

High expression Lola = black 


```{r Day 0 all mods of interest KEGG}

Day0_WGCNA_sigmodules <- as.data.frame(c('turquoise', 
                                         'blue', 
                                         'yellow', 
                                         'brown', 
                                         'black'))
# prep loop for cumulative output table 
df_total             <- data.frame() # start dataframe 
KEGG.Day0           <- data.frame(matrix(nrow = 1, ncol = 10)) # create dataframe to save cumunalitively during for loop
colnames(KEGG.Day0) <- c('Day', 'modColor', 'KEGGID_pathway', 'pathway.name' , 
                         'Count.enriched', 'Count.per.pathway',  
                         'Gene.IDs', 'p.value', 'log10_pvalue', 'Rich_factor') # names for comuns in the for loop

for (i in 1:nrow(Day0_WGCNA_sigmodules)) {
    modColor         <- Day0_WGCNA_sigmodules[i,1]
    loopmodColor_cor <- paste("MM.", modColor, sep = '')
    loopmodColor_p   <- paste("p.MM.", modColor, sep = '')
    ModuleLoop       <- as.data.frame(Day0_ModuleMembership %>% 
                                        dplyr::filter(moduleColor %in% modColor)  %>% 
                                        dplyr::select(c('geneSymbol', 
                                                        'KEGG_ID', 
                                                        'moduleColor', 
                                                        loopmodColor_p, 
                                                        loopmodColor_cor)) %>% 
                                                      #dplyr::filter(.[[5]] < 0.05 & .[[6]] > 0.6) %>% # filter based on thresholds set in the WGCNA + DESeq2 overlap w/PCA 
                                        dplyr::select(c('geneSymbol', 'KEGG_ID',4)) %>% 
                                        na.omit() %>% 
                                        dplyr::mutate(KEGG_ID = gsub(".*:","",KEGG_ID)) %>% 
                                        unnest(KEGG_ID))
    geneList <- ModuleLoop[,3]
    names(geneList) <- ModuleLoop$KEGG_ID
    # Wilcoxon test for each pathway
    pVals.by.pathway <- t(sapply(names(genes.by.pathway),
                                 function(pathway) {
                                   pathway.genes <- genes.by.pathway[[pathway]]
                                   list.genes.in.pathway <- intersect(names(geneList), pathway.genes)
                                   list.genes.not.in.pathway <- setdiff(names(geneList), list.genes.in.pathway)
                                   scores.in.pathway <- geneList[list.genes.in.pathway]
                                   scores.not.in.pathway <- geneList[list.genes.not.in.pathway]
                                   if (length(scores.in.pathway) > 0){
                                     p.value <- wilcox.test(scores.in.pathway, scores.not.in.pathway, alternative = "less")$p.value
                                   } else{
                                     p.value <- NA
                                   }
                                   return(c(p.value = p.value, Annotated = length(list.genes.in.pathway), GeneIDs = list(list.genes.in.pathway) ))
                                 }
    ))
    # Assemble output table
    outdat <- data.frame(pathway.code = rownames(pVals.by.pathway)) %>% 
      dplyr::mutate(Count.enriched    = as.numeric(as.data.frame(pVals.by.pathway)$Annotated), 
                    Count.per.pathway = (Genes_per_pathway %>%  
                                               dplyr::filter(KEGGID_pathway == rownames(pVals.by.pathway)))$Num.genes.per.pathway,
                    Gene.IDs = as.data.frame(pVals.by.pathway)$GeneIDs) %>% 
      dplyr::rename(KEGGID_pathway = pathway.code) %>% 
      dplyr::mutate(pathway.name = as.character((pathways.list)[pathway.code]), 
                    pathway.name = gsub(" -.*","",pathway.name), # remove ' - Crassostrea gigas (Pacific oyster)'
                    p.value =  pVals.by.pathway[,"p.value"], 
                    Day = 'Day0', 
                    modColor = modColor) %>% 
      dplyr::filter(p.value < 0.05) %>% 
      na.omit() %>% 
      dplyr::mutate(log10_pvalue = -log10(as.numeric(p.value)),
                    Rich_factor = as.numeric(Count.enriched) / as.numeric(Count.per.pathway))
                    
    
    KEGG.Day0 <- rbind(KEGG.Day0,outdat) #bind to a cumulative list dataframe
    print(KEGG.Day0) # print to monitor progress
}
View(KEGG.Day0)


# KEGG Rich factor plot 
D0_KEGG <- KEGG.Day0 %>%  
              na.omit() %>% # omit 1 row with NAs
              ggplot(aes(x=reorder(pathway.name, as.numeric(Rich_factor)), y= as.numeric(Rich_factor) )) + 
                geom_point( aes(col=modColor, size=Count.enriched)) +   # Draw points
                geom_segment(aes(x=pathway.name, 
                                 xend=pathway.name, 
                                 y=min(as.numeric(Rich_factor)), 
                                 yend=max(as.numeric(Rich_factor))),  
                             linetype=NA, 
                             size=0) +   # Draw dashed lines
                labs(title="Day 0", 
                     x = "Pathway",
                     y = "Rich Factor") +
                coord_flip() +
                theme_bw() +
                facet_wrap(~modColor, scales = 'free_y')

pdf(paste("Output/Day0/KEGG/RichFactorPlot.pdf", sep =''), 
    width=10, 
    height=9)
print(D0_KEGG)
dev.off()
            

```

# DAY 1 DATA - FOR LOOP 

## About:

Day 1 (after several days of recovery)

High expression Fed = greenyellow  

High expression Starved = green 

High expression Wild =  brown 

High expression Lola = blue 

```{r Day 1 all mods of interest KEGG}

Day1_WGCNA_sigmodules <- as.data.frame(c('greenyellow', 
                                         'green', 
                                         'brown', 
                                         'blue'))
# prep loop for cumulative output table 
df_total             <- data.frame() # start dataframe 
KEGG.Day1            <- data.frame(matrix(nrow = 1, ncol = 10)) # create dataframe to save cumunalitively during for loop
colnames(KEGG.Day1) <- c('Day', 'modColor', 'KEGGID_pathway', 'pathway.name' , 
                         'Count.enriched', 'Count.per.pathway',  
                         'Gene.IDs', 'p.value', 'log10_pvalue', 'Rich_factor') # names for comuns in the for loop

for (i in 1:nrow(Day1_WGCNA_sigmodules)) {
    modColor         <- Day1_WGCNA_sigmodules[i,1]
    loopmodColor_cor <- paste("MM.", modColor, sep = '')
    loopmodColor_p   <- paste("p.MM.", modColor, sep = '')
    ModuleLoop       <- as.data.frame(Day1_ModuleMembership %>% 
                                        dplyr::filter(moduleColor %in% modColor)  %>% 
                                        dplyr::select(c('geneSymbol', 
                                                        'KEGG_ID', 
                                                        'geneSymbol', 
                                                        'moduleColor', 
                                                        loopmodColor_p, 
                                                        loopmodColor_cor)) %>% 
                                                      #dplyr::filter(.[[5]] < 0.05 & .[[6]] > 0.6) %>% # filter based on thresholds set in the WGCNA + DESeq2 overlap w/PCA 
                                        dplyr::select(c('geneSymbol', 'KEGG_ID',5)) %>% 
                                        na.omit() %>% 
                                        dplyr::mutate(KEGG_ID = gsub(".*:","",KEGG_ID)) %>% 
                                        unnest(KEGG_ID))
    geneList <- ModuleLoop[,3]
    names(geneList) <- ModuleLoop$KEGG_ID
    # Wilcoxon test for each pathway
    pVals.by.pathway <- t(sapply(names(genes.by.pathway),
                                 function(pathway) {
                                   pathway.genes <- genes.by.pathway[[pathway]]
                                   list.genes.in.pathway <- intersect(names(geneList), pathway.genes)
                                   list.genes.not.in.pathway <- setdiff(names(geneList), list.genes.in.pathway)
                                   scores.in.pathway <- geneList[list.genes.in.pathway]
                                   scores.not.in.pathway <- geneList[list.genes.not.in.pathway]
                                   if (length(scores.in.pathway) > 0){
                                     p.value <- wilcox.test(scores.in.pathway, scores.not.in.pathway, alternative = "less")$p.value
                                   } else{
                                     p.value <- NA
                                   }
                                   return(c(p.value = p.value, Annotated = length(list.genes.in.pathway), GeneIDs = list(list.genes.in.pathway) ))
                                 }
    ))
    # Assemble output table
    outdat <- data.frame(pathway.code = rownames(pVals.by.pathway)) %>% 
      dplyr::mutate(Count.enriched    = as.numeric(as.data.frame(pVals.by.pathway)$Annotated), 
                    Count.per.pathway = (Genes_per_pathway %>%  
                                               dplyr::filter(KEGGID_pathway == rownames(pVals.by.pathway)))$Num.genes.per.pathway,
                    Gene.IDs = as.data.frame(pVals.by.pathway)$GeneIDs) %>% 
      dplyr::rename(KEGGID_pathway = pathway.code) %>% 
      dplyr::mutate(pathway.name = as.character((pathways.list)[pathway.code]), 
                    pathway.name = gsub(" -.*","",pathway.name), # remove ' - Crassostrea gigas (Pacific oyster)'
                    p.value =  pVals.by.pathway[,"p.value"], 
                    Day = 'Day1', 
                    modColor = modColor) %>% 
      dplyr::filter(p.value < 0.05) %>% 
      na.omit() %>% 
      dplyr::mutate(log10_pvalue = -log10(as.numeric(p.value)),
                    Rich_factor = as.numeric(Count.enriched) / as.numeric(Count.per.pathway))
    
    KEGG.Day1 <- rbind(KEGG.Day1,outdat) #bind to a cumulative list dataframe
    print(KEGG.Day1) # print to monitor progress
}
View(KEGG.Day1)

D1_KEGG<- KEGG.Day1 %>%  
              na.omit() %>% # omit 1 row with NAs
              ggplot(aes(x=reorder(pathway.name, as.numeric(Rich_factor)), y= as.numeric(Rich_factor) )) + 
                geom_point( aes(col=modColor, size=Count.enriched)) +   # Draw points
                geom_segment(aes(x=pathway.name, 
                                 xend=pathway.name, 
                                 y=min(as.numeric(Rich_factor)), 
                                 yend=max(as.numeric(Rich_factor))),  
                             linetype=NA, 
                             size=0) +   # Draw dashed lines
                labs(title="Day 1", 
                     x = "Pathway",
                     y = "Rich Factor") +
                coord_flip() +
                theme_bw() +
                facet_wrap(~modColor, scales = 'free_y')


pdf(paste("Output/Day1/KEGG/RichFactorPlot.pdf", sep =''), 
    width=10, 
    height=9)
print(D1_KEGG)
dev.off()
            


```

High expression Fed = greenyellow  **compare to Turquoise day 0**

```{r Overlapped responses - High expression when fed}

# High expression when Fed

# merge genes in greenyellow that overlap with turquoise day 0
Day1_greenyellow <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'greenyellow')
Day0_turquoise   <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'turquoise')
HighExp_Fed      <- Day1_greenyellow %>% na.omit() %>% filter(geneSymbol %in% Day0_turquoise$geneSymbol)
nrow(HighExp_Fed)
HighExp_Fed_formatted  <- as.data.frame(HighExp_Fed %>% 
                                  dplyr::select(c('geneSymbol', 
                                                  'KEGG_ID', 
                                                  'moduleColor', 
                                                   p.MM.greenyellow, 
                                                   p.MM.greenyellow)) %>% 
                                        dplyr::select(c('geneSymbol', 'KEGG_ID',4)) %>% 
                                        na.omit() %>% 
                                        dplyr::mutate(KEGG_ID = gsub(".*:","",KEGG_ID)) %>% 
                                        unnest(KEGG_ID))

geneList        <- HighExp_Fed_formatted[,3]
names(geneList) <- HighExp_Fed_formatted$KEGG_ID
# Wilcoxon test for each pathway
pVals.by.pathway <- t(sapply(names(genes.by.pathway), 
                                 function(pathway) {
                                   pathway.genes <- genes.by.pathway[[pathway]]
                                   list.genes.in.pathway <- intersect(names(geneList), pathway.genes)
                                   list.genes.not.in.pathway <- setdiff(names(geneList), list.genes.in.pathway)
                                   scores.in.pathway <- geneList[list.genes.in.pathway]
                                   scores.not.in.pathway <- geneList[list.genes.not.in.pathway]
                                   if (length(scores.in.pathway) > 0){
                                     p.value <- wilcox.test(scores.in.pathway, 
                                                            scores.not.in.pathway, 
                                                            alternative = "less")$p.value
                                   } else{
                                     p.value <- NA
                                   }
                                   return(c(p.value = p.value, 
                                            Annotated = length(list.genes.in.pathway), 
                                            GeneIDs = list(list.genes.in.pathway) ))
                                 }))
# Assemble output table
HighExp_Fed_KEGG <- data.frame(pathway.code = rownames(pVals.by.pathway)) %>% 
      dplyr::mutate(Count.enriched    = as.numeric(as.data.frame(pVals.by.pathway)$Annotated), 
                    Count.per.pathway = (Genes_per_pathway %>%  
                                               dplyr::filter(KEGGID_pathway == rownames(pVals.by.pathway)))$Num.genes.per.pathway,
                    Gene.IDs = as.data.frame(pVals.by.pathway)$GeneIDs) %>% 
      dplyr::rename(KEGGID_pathway = pathway.code) %>% 
      dplyr::mutate(pathway.name = as.character((pathways.list)[pathway.code]), 
                    pathway.name = gsub(" -.*","",pathway.name), # remove ' - Crassostrea gigas (Pacific oyster)'
                    p.value =  pVals.by.pathway[,"p.value"], 
                    Day = 'Day1', 
                    modColor = modColor) %>% 
      dplyr::filter(p.value < 0.05) %>% 
      na.omit() %>% 
      dplyr::mutate(log10_pvalue = -log10(as.numeric(p.value)),
                    Rich_factor = as.numeric(Count.enriched) / as.numeric(Count.per.pathway))
HighExp_Fed_KEGG

Fed_KEGG<- HighExp_Fed_KEGG %>%  
              na.omit() %>% # omit 1 row with NAs
              ggplot(aes(x=reorder(pathway.name, as.numeric(Rich_factor)), y= as.numeric(Rich_factor) )) + 
                geom_point( aes(col="day 0 turquoise  & day 1 greenyellow", size=Count.enriched)) +   # Draw points
                geom_segment(aes(x=pathway.name, 
                                 xend=pathway.name, 
                                 y=min(as.numeric(Rich_factor)), 
                                 yend=max(as.numeric(Rich_factor))),  
                             linetype=NA, 
                             size=0) +   # Draw dashed lines
                labs(title=paste0("Fed = High Expression ",nrow(HighExp_Fed), " genes"), 
                     x = "Pathway",
                     y = "Rich Factor") +
                coord_flip() +
                theme_bw() 


pdf(paste("Output/KEGG_RichFactor_Fed.pdf", sep =''), 
    width=10, 
    height=9)
print(Fed_KEGG)
dev.off()

```
High expression Starved = Day 1 green  **compare to Day 0 blue**

```{r Overlapped responses - High expression when STARVED}

# High expression when Fed

# merge genes in greenyellow that overlap with turquoise day 0
Day1_green      <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'green')
Day0_blue       <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'blue')
HighExp_Starved <- Day1_green %>% na.omit() %>% filter(geneSymbol %in% Day0_blue$geneSymbol)
nrow(HighExp_Starved)
HighExp_Starved_formatted  <- as.data.frame(HighExp_Starved %>% 
                                  dplyr::select(c('geneSymbol', 
                                                  'KEGG_ID', 
                                                  'moduleColor', 
                                                   p.MM.green, 
                                                   p.MM.green)) %>% 
                                        dplyr::select(c('geneSymbol', 'KEGG_ID',4)) %>% 
                                        na.omit() %>% 
                                        dplyr::mutate(KEGG_ID = gsub(".*:","",KEGG_ID)) %>% 
                                        unnest(KEGG_ID))

geneList        <- HighExp_Starved_formatted[,3]
names(geneList) <- HighExp_Starved_formatted$KEGG_ID
# Wilcoxon test for each pathway
pVals.by.pathway <- t(sapply(names(genes.by.pathway), 
                                 function(pathway) {
                                   pathway.genes <- genes.by.pathway[[pathway]]
                                   list.genes.in.pathway <- intersect(names(geneList), pathway.genes)
                                   list.genes.not.in.pathway <- setdiff(names(geneList), list.genes.in.pathway)
                                   scores.in.pathway <- geneList[list.genes.in.pathway]
                                   scores.not.in.pathway <- geneList[list.genes.not.in.pathway]
                                   if (length(scores.in.pathway) > 0){
                                     p.value <- wilcox.test(scores.in.pathway, 
                                                            scores.not.in.pathway, 
                                                            alternative = "less")$p.value
                                   } else{
                                     p.value <- NA
                                   }
                                   return(c(p.value = p.value, 
                                            Annotated = length(list.genes.in.pathway), 
                                            GeneIDs = list(list.genes.in.pathway) ))
                                 }))
# Assemble output table
HighExp_Starved_KEGG <- data.frame(pathway.code = rownames(pVals.by.pathway)) %>% 
      dplyr::mutate(Count.enriched    = as.numeric(as.data.frame(pVals.by.pathway)$Annotated), 
                    Count.per.pathway = (Genes_per_pathway %>%  
                                               dplyr::filter(KEGGID_pathway == rownames(pVals.by.pathway)))$Num.genes.per.pathway,
                    Gene.IDs = as.data.frame(pVals.by.pathway)$GeneIDs) %>% 
      dplyr::rename(KEGGID_pathway = pathway.code) %>% 
      dplyr::mutate(pathway.name = as.character((pathways.list)[pathway.code]), 
                    pathway.name = gsub(" -.*","",pathway.name), # remove ' - Crassostrea gigas (Pacific oyster)'
                    p.value =  pVals.by.pathway[,"p.value"], 
                    Day = 'Day1', 
                    modColor = modColor) %>% 
      dplyr::filter(p.value < 0.05) %>% 
      na.omit() %>% 
      dplyr::mutate(log10_pvalue = -log10(as.numeric(p.value)),
                    Rich_factor = as.numeric(Count.enriched) / as.numeric(Count.per.pathway))
HighExp_Starved_KEGG

Starved_KEGG<- HighExp_Starved_KEGG %>%  
              na.omit() %>% # omit 1 row with NAs
              ggplot(aes(x=reorder(pathway.name, as.numeric(Rich_factor)), y= as.numeric(Rich_factor) )) + 
                geom_point( aes(col="day 0 blue  & day 1 green", size=Count.enriched)) +   # Draw points
                geom_segment(aes(x=pathway.name, 
                                 xend=pathway.name, 
                                 y=min(as.numeric(Rich_factor)), 
                                 yend=max(as.numeric(Rich_factor))),  
                             linetype=NA, 
                             size=0) +   # Draw dashed lines
                labs(title=paste0("Starved = High Expression ",nrow(HighExp_Starved), " genes"), 
                     x = "Pathway",
                     y = "Rich Factor") +
                coord_flip() +
                theme_bw() 


pdf(paste("Output/KEGG_RichFactor_Starved.pdf", sep =''), 
    width=10, 
    height=9)
print(Starved_KEGG)
dev.off()

```


High expression Wild =  day 1 brown **comapre to day 0 yellow**

```{r Overlapped responses - High expression for WILD population}

# High expression when Fed

# merge genes in greenyellow that overlap with turquoise day 0
Day1_brown     <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'brown')
Day0_yellow    <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'yellow')
HighExp_Wild   <- Day1_brown %>% na.omit() %>% filter(geneSymbol %in% Day0_yellow$geneSymbol)
nrow(HighExp_Wild)
HighExp_Wild_formatted  <- as.data.frame(HighExp_Wild %>% 
                                  dplyr::select(c('geneSymbol', 
                                                  'KEGG_ID', 
                                                  'moduleColor', 
                                                   p.MM.brown, 
                                                   p.MM.brown)) %>% 
                                        dplyr::select(c('geneSymbol', 'KEGG_ID',4)) %>% 
                                        na.omit() %>% 
                                        dplyr::mutate(KEGG_ID = gsub(".*:","",KEGG_ID)) %>% 
                                        unnest(KEGG_ID))

geneList        <- HighExp_Wild_formatted[,3]
names(geneList) <- HighExp_Wild_formatted$KEGG_ID
# Wilcoxon test for each pathway
pVals.by.pathway <- t(sapply(names(genes.by.pathway), 
                                 function(pathway) {
                                   pathway.genes <- genes.by.pathway[[pathway]]
                                   list.genes.in.pathway <- intersect(names(geneList), pathway.genes)
                                   list.genes.not.in.pathway <- setdiff(names(geneList), list.genes.in.pathway)
                                   scores.in.pathway <- geneList[list.genes.in.pathway]
                                   scores.not.in.pathway <- geneList[list.genes.not.in.pathway]
                                   if (length(scores.in.pathway) > 0){
                                     p.value <- wilcox.test(scores.in.pathway, 
                                                            scores.not.in.pathway, 
                                                            alternative = "less")$p.value
                                   } else{
                                     p.value <- NA
                                   }
                                   return(c(p.value = p.value, 
                                            Annotated = length(list.genes.in.pathway), 
                                            GeneIDs = list(list.genes.in.pathway) ))
                                 }))
# Assemble output table
HighExp_Wild_KEGG <- data.frame(pathway.code = rownames(pVals.by.pathway)) %>% 
      dplyr::mutate(Count.enriched    = as.numeric(as.data.frame(pVals.by.pathway)$Annotated), 
                    Count.per.pathway = (Genes_per_pathway %>%  
                                               dplyr::filter(KEGGID_pathway == rownames(pVals.by.pathway)))$Num.genes.per.pathway,
                    Gene.IDs = as.data.frame(pVals.by.pathway)$GeneIDs) %>% 
      dplyr::rename(KEGGID_pathway = pathway.code) %>% 
      dplyr::mutate(pathway.name = as.character((pathways.list)[pathway.code]), 
                    pathway.name = gsub(" -.*","",pathway.name), # remove ' - Crassostrea gigas (Pacific oyster)'
                    p.value =  pVals.by.pathway[,"p.value"], 
                    Day = 'Day1', 
                    modColor = modColor) %>% 
      dplyr::filter(p.value < 0.05) %>% 
      na.omit() %>% 
      dplyr::mutate(log10_pvalue = -log10(as.numeric(p.value)),
                    Rich_factor = as.numeric(Count.enriched) / as.numeric(Count.per.pathway))
HighExp_Wild_KEGG

Wild_KEGG<- HighExp_Wild_KEGG %>%  
              na.omit() %>% # omit 1 row with NAs
              ggplot(aes(x=reorder(pathway.name, as.numeric(Rich_factor)), y= as.numeric(Rich_factor) )) + 
                geom_point( aes(col="day 0 yellow  & day 1 brown", size=Count.enriched)) +   # Draw points
                geom_segment(aes(x=pathway.name, 
                                 xend=pathway.name, 
                                 y=min(as.numeric(Rich_factor)), 
                                 yend=max(as.numeric(Rich_factor))),  
                             linetype=NA, 
                             size=0) +   # Draw dashed lines
                labs(title=paste0("Wild = High Expression ",nrow(HighExp_Wild), " genes"), 
                     x = "Pathway",
                     y = "Rich Factor") +
                coord_flip() +
                theme_bw() 


pdf(paste("Output/KEGG_RichFactor_Wild.pdf", sep =''), 
    width=10, 
    height=9)
print(Wild_KEGG)
dev.off()

```

High expression Lola = day 1 blue - **compare to day 0 black**

```{r Overlapped responses - High expression for Lola population}

# High expression when Fed

# merge genes in greenyellow that overlap with turquoise day 0
Day1_blue      <- Day1_ModuleMembership %>% dplyr::filter(moduleColor %in% 'blue')
Day0_black     <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'black')
HighExp_Lola   <- Day1_blue %>% na.omit() %>% filter(geneSymbol %in% Day0_black$geneSymbol)
nrow(HighExp_Lola)
HighExp_Lola_formatted  <- as.data.frame(HighExp_Lola %>% 
                                  dplyr::select(c('geneSymbol', 
                                                  'KEGG_ID', 
                                                  'moduleColor', 
                                                   p.MM.blue, 
                                                   p.MM.blue)) %>% 
                                        dplyr::select(c('geneSymbol', 'KEGG_ID',4)) %>% 
                                        na.omit() %>% 
                                        dplyr::mutate(KEGG_ID = gsub(".*:","",KEGG_ID)) %>% 
                                        unnest(KEGG_ID))

geneList        <- HighExp_Lola_formatted[,3]
names(geneList) <- HighExp_Lola_formatted$KEGG_ID
# Wilcoxon test for each pathway
pVals.by.pathway <- t(sapply(names(genes.by.pathway), 
                                 function(pathway) {
                                   pathway.genes <- genes.by.pathway[[pathway]]
                                   list.genes.in.pathway <- intersect(names(geneList), pathway.genes)
                                   list.genes.not.in.pathway <- setdiff(names(geneList), list.genes.in.pathway)
                                   scores.in.pathway <- geneList[list.genes.in.pathway]
                                   scores.not.in.pathway <- geneList[list.genes.not.in.pathway]
                                   if (length(scores.in.pathway) > 0){
                                     p.value <- wilcox.test(scores.in.pathway, 
                                                            scores.not.in.pathway, 
                                                            alternative = "less")$p.value
                                   } else{
                                     p.value <- NA
                                   }
                                   return(c(p.value = p.value, 
                                            Annotated = length(list.genes.in.pathway), 
                                            GeneIDs = list(list.genes.in.pathway) ))
                                 }))
# Assemble output table
HighExp_Lola_KEGG <- data.frame(pathway.code = rownames(pVals.by.pathway)) %>% 
      dplyr::mutate(Count.enriched    = as.numeric(as.data.frame(pVals.by.pathway)$Annotated), 
                    Count.per.pathway = (Genes_per_pathway %>%  
                                               dplyr::filter(KEGGID_pathway == rownames(pVals.by.pathway)))$Num.genes.per.pathway,
                    Gene.IDs = as.data.frame(pVals.by.pathway)$GeneIDs) %>% 
      dplyr::rename(KEGGID_pathway = pathway.code) %>% 
      dplyr::mutate(pathway.name = as.character((pathways.list)[pathway.code]), 
                    pathway.name = gsub(" -.*","",pathway.name), # remove ' - Crassostrea gigas (Pacific oyster)'
                    p.value =  pVals.by.pathway[,"p.value"], 
                    Day = 'Day1', 
                    modColor = modColor) %>% 
      dplyr::filter(p.value < 0.05) %>% 
      na.omit() %>% 
      dplyr::mutate(log10_pvalue = -log10(as.numeric(p.value)),
                    Rich_factor = as.numeric(Count.enriched) / as.numeric(Count.per.pathway))
HighExp_Lola_KEGG

Lola_KEGG<- HighExp_Lola_KEGG %>%  
              na.omit() %>% # omit 1 row with NAs
              ggplot(aes(x=reorder(pathway.name, as.numeric(Rich_factor)), y= as.numeric(Rich_factor) )) + 
                geom_point( aes(col="day 0 black  & day 1 blue", size=Count.enriched)) +   # Draw points
                geom_segment(aes(x=pathway.name, 
                                 xend=pathway.name, 
                                 y=min(as.numeric(Rich_factor)), 
                                 yend=max(as.numeric(Rich_factor))),  
                             linetype=NA, 
                             size=0) +   # Draw dashed lines
                labs(title=paste0("Lola = High Expression ",nrow(HighExp_Lola), " genes"), 
                     x = "Pathway",
                     y = "Rich Factor") +
                coord_flip() +
                theme_bw() 


pdf(paste("Output/KEGG_RichFactor_Lola.pdf", sep =''), 
    width=10, 
    height=9)
print(Lola_KEGG)
dev.off()

```



