---
title: "WGCNA_Cvirginica_Feed_Priming"
author: "Samuel Gurr"
date: "2023-08-31"
output: html_document
---

```{r setup, include=FALSE}
# load libraries - notes show the install command needed to install (pre installed)
library(goseq)
library(dplyr)
library(forcats)
library(ggplot2)
library(gridExtra)
library(tidyr)
library(grDevices)
library(reshape2)
library(Rmisc)
library(ggpubr)
library(tibble)
library(hrbrthemes)
library(gridExtra)
library(tidyr)
library(zoo)
library(ComplexHeatmap)
library(circlize)
library(GSEABase)
library(data.table)
library(stringr)
# SET WORKING DIRECTORY
knitr::opts_knit$set(root.dir = "C:/Users/samjg/Documents/Github_repositories/Cvirginica_KM_paper/RAnalysis") # sets the working 
```


# LOAD THE MODULE-MEMBERSHIP MASTER TABLES 
```{r} 


Day0_ModuleMembership  <- read.csv(file="Output/Day0/Day0.WGCNA_ModulMembership.csv", sep=',', header=TRUE) %>% 
                                       dplyr::select(c('X','GO.terms', 'KEGG_ID', 'geneSymbol', 'Protein_name','moduleColor')) %>%  
                                       na.omit() %>% 
                                       dplyr::rename(Cvirginica_GeneID = 'X') %>% 
                                       dplyr::mutate(Day = "Day0")



Day1_ModuleMembership  <- read.csv(file="Output/Day1/Day1.WGCNA_ModulMembership.csv", sep=',', header=TRUE) %>% 
                                       dplyr::select(c('X','GO.terms', 'KEGG_ID', 'geneSymbol', 'Protein_name','moduleColor')) %>%  
                                       na.omit() %>% 
                                       dplyr::rename(Cvirginica_GeneID = 'X') %>% 
                                       dplyr::mutate(Day = "Day1")

```

# What are we ding here? 

* we found that Fed and Starved oysters(regardled of wild or lola) had high transcript abundnace via WGCNA at termination of the starvation trip 
* (1) Are these genes continuously expressed at that level, *even after several days of recovery (high food)?* 
* (2) Or... do these genes approximate the transcript abundnace by the fed oysters - those that did not experience a starvation event
  * Transcripts in number (1) represent a legacy effect of starvation, those severly affected by food scarcity and unresponsive to food supply
  * Transcripts in number (2) represent responsive functions - whether food supply permits more (otherwise not prioritized during starvation) or less (recovery response) transcript abundance 
  
# Timepoint 3 (as 'day0' in these scripts) is the end of the starvation challenge
  - turquise == suppressed transcripts under starvation (relative to fed)
  - blue     == high trnacript abundnace under starvation (relative to fed)


## Objectives: isolate genes that responded to food from the master WGCNA file on timepoint 3 (day0)
and overlay them into the subseqent timepoint, quantify the number present and in which modules they are represented
```{r less exploreshall we!}

# filter the master file of the first timepoint
Starved_HighExp_T1 <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'blue')
Starved_LowExp_T1  <- Day0_ModuleMembership %>% dplyr::filter(moduleColor %in% 'turquoise')

# we just need the unique idnetifier - lets do gene name

Starved_HighExp_ID <- Starved_HighExp_T1[,1]
length(Starved_HighExp_ID) # 4544 - number of genes! damn thats a lot..

Starved_LowExp_ID  <- Starved_LowExp_T1[,1]
length(Starved_LowExp_ID) # 5236 - number of genes! damn thats a lot..

# use these geneIDs to filter the subseqent timepoint - quanitfy the number of genes and which models they are in
Starved_HighExp_T2 <-  Day1_ModuleMembership %>% dplyr::filter(Cvirginica_GeneID %in% Starved_HighExp_T1$Cvirginica_GeneID)
nrow(Starved_HighExp_T1) - nrow(Starved_HighExp_T2) # 632 are not present in the WGCNA data for the subseqent timepoint
100 - (  ((nrow(Starved_HighExp_T1) - nrow(Starved_HighExp_T2)) / nrow(Starved_HighExp_T1) )*100 ) # 86.09155 found in T2!



Starved_LowExp_T2 <-  Day1_ModuleMembership %>% dplyr::filter(Cvirginica_GeneID %in% Starved_LowExp_T1$Cvirginica_GeneID)
nrow(Starved_LowExp_T1) - nrow(Starved_LowExp_T2) # 143 are not present in the WGCNA data for the subseqent timepoint
100 - (  ((nrow(Starved_LowExp_T1) - nrow(Starved_LowExp_T2)) / nrow(Starved_LowExp_T1) )*100 ) # 97.26891 found in T2!



# First, how many genes were in these modules in T2 ('Day1') to start?
T2_summary <- as.data.frame(
                    Day1_ModuleMembership %>% dplyr::select(moduleColor ,Cvirginica_GeneID) %>% 
                          na.omit() %>% 
                          dplyr::group_by(moduleColor) %>% 
                          dplyr::summarise(n= n())  )


# High Expression starved - overlay into the T2_summary to see where these genes are expressed!
T2_StarvationHighExp <- as.data.frame( 
                  Starved_HighExp_T2 %>% dplyr::select(moduleColor, Cvirginica_GeneID) %>% 
                          na.omit() %>% 
                          dplyr::group_by(moduleColor) %>% 
                          dplyr::summarise(n_HighExp_Starved= n())  )

Starvation_HighExp_Prop_T2 <- merge(T2_summary, T2_StarvationHighExp, by ="moduleColor") %>% dplyr::mutate(Percent_HighExpStarved = (n_HighExp_Starved / (n+n_HighExp_Starved))*100 )
Starvation_HighExp_Prop_T2

# High Expression starved - overlay into the T2_summary to see where these genes are expressed!
T2_StarvationLowExp <- as.data.frame( 
                  Starved_LowExp_T2 %>% dplyr::select(moduleColor, Cvirginica_GeneID) %>% 
                          na.omit() %>% 
                          dplyr::group_by(moduleColor) %>% 
                          dplyr::summarise(n_LowExp_Starved= n())  )

Starvation_LowExp_Prop_T2 <- merge(T2_summary, T2_StarvationLowExp, by ="moduleColor") %>% dplyr::mutate(Percent_LowExpStarved = (n_LowExp_Starved / (n+n_LowExp_Starved))*100 )
Starvation_LowExp_Prop_T2


# Although no directly clear evidence of module-specific carryoer here. we do have something intersting! 
# High Exp genes under starvation at timepoint1 have the highest proportion of module 'cyan' (34.8%) - the low Exp genes under starvation make up the lowest propotion of module 'cyan' (17.8%)
# module cyan is uniqqly associated wit high expession of oysters that were starved! 



View(Starved_HighExp_T2 %>% dplyr::filter(moduleColor %in% 'cyan') %>% dplyr::select(Protein_name) %>% dplyr::arrange(Protein_name))
```